\Large \bf Cyclone: Warp Speed Replication for Persistent Key Value Stores

### 2017-07-27 paper/paper.tex
Updates.

\title{\Large \bf Cyclone: Warp Speed Replication for Key Value Stores}

@@

The first key contribution of Cyclone is in showing how a small amount of
non-volatile memory can be combined with the flash SSD based persistent logs
when providing high availability.  The primary cause of slowdown when turning on
a write ahead log on a single machine in Figure~\ref{fig:problem} is the need to
do synchronous block IO for every update. For evenmoderately sized updates, this
leads to severe write amplification as a 4KB page must be synchronously written
to the log file on flash. Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are expensive - retailing at approximately 2X the price of a standard DIMM [find
  ref]. This means that they will likely be available in small quantities. We
compensate for this by periodically draining the NVDIMM log into a log placed on
a standard SSD on the system.

The second key contribution of this work is in showing that the performance
overheads in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum - a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory eliminating block IO from the persistence step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. On
40Gbps infiniband, Mojim reported being able to mirror data at 4GB/s. In
contrast, Cyclone can replicate data three ways using 4*10 Gbps commodity
ethernet NICs at a peak of roughly 3.7 GB/s. Notably this is while Cyclone is
also running the RAFT consensus protocol and draining the log to a high capacity
flash SSD, while Mojim is doing simple mirroring of main memory. One can
conclude that network offload via RDMA and Infiniband is not an a-priori
requirement for high throughput replication. This is made possible by Cyclone's
approach that relieves the CPU from the burden of data copies. Of course, the
latency of a hop on commodity ethernet at 5 us with our DPDK stack is larger
than the close to 2us latency with Infiniband. Depending on the use case for the
key value store application the extra 6us with ethernet for the round trip to
the replica might be a concern for some applications. We note that mapping the
commutativity of key value store interfaces to multiple logs in Cyclone is a
technique that can also be applied to RDMA. The authors of Mojim confirm that
concurrency is a significant determinant of performance when replicating over
RDMA.

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency numbers are
excellent due to the FPGA that cuts out the path from CPU caches to the NIC and
are reported to be as low as 3us for a round trip from leader to follower
replica compared to the 6us we observe with DPDK. Notably however they reported
7us when using TCP/IP from the FPGA rather than their own connection oriented
network protocol. On the other hand their peak replication throughput is 6M
replications/sec on 66 Gbps of aggregate network bandwith, comparable to the
same mark obtained by Cyclone on an aggregate 40 Gbps of network bandwidth. We
believe the general applicability afforded by our software only solution
together with the fact that we provide persistence by design makes it a
compelling alternative to an FPGA based solution even given the larger latency.

@@

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request is currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores, along the lines of similar work~\cite{raft_lin}, using a fixed client
space and maintaining information about last seen sequence numbers from clients
in NVM. Distributed transactions on Cyclone shards however are outside the scope
of this paper.

>>>>

\title{\Large \bf Cyclone: Warp Speed Replication for Persistent Key Value Stores}

@@

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update. For small to moderately sized updates, this leads to severe
write amplification as a 4KB page must be synchronously written to the log file
on flash. Cyclone avoids this by making use of directly attached persistent
memory on the server. This can be in the form of NVDIMMs (DIMMs with an added
ultracapacitor to dump state to a small amount of attached flash) or newer forms
of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs are expensive
- retailing at approximately 2X the price of a standard DIMM [find ref]. This
means that they will likely be available in small quantities. We compensate for
this by periodically draining the NVDIMM log into a log placed on a standard SSD
on the system.

The second key contribution of this work is in showing that the throughput
overheads in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum - a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory eliminating block IO from the persistence step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. A
comparison is hard since Mojim is designed for the hypothetical scenario where
there is enough persistent memory to hold all persistent data, while Cyclone is
designed for the practical case where both the write-ahead log and key value
store data structures need to be maintained largely on flash. Nevertheless we
can compare their replication performance for the case where we use only the NVM
log in Cyclone. On 40Gbps infiniband, Mojim reported being able to mirror data
at 4GB/s. In contrast, Cyclone can replicate data three ways using 4*10 Gbps
commodity ethernet NICs at a peak of roughly 3.7 GB/s. Notably this is while
Cyclone is also running the RAFT consensus protocol, while Mojim is doing simple
mirroring of main memory. One can conclude that network offload via RDMA and
Infiniband is not an a-priori requirement for high throughput replication. This
is made possible by Cyclone's approach that relieves the CPU from the burden of
data copies. Of course, the latency of a hop on commodity ethernet at 5 us with
our DPDK stack is larger than the close to 2us latency with
Infiniband. Depending on the use case for the key value store application the
extra 6us with ethernet for the round trip to the replica might be a concern for
some applications. We note that mapping the commutativity of key value store
interfaces to multiple logs in Cyclone is a technique that can also be applied
to RDMA. The authors of Mojim confirm that concurrency is a significant
determinant of performance when replicating over RDMA.

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency numbers are
excellent due to the FPGA that cuts out the path from CPU caches to the NIC and
are reported to be as low as 3us for a round trip from leader to follower
replica compared to the 6us we observe with DPDK. Notably however they reported
7us when using TCP/IP from the FPGA rather than their own connection oriented
network protocol. On the other hand their peak replication throughput is 6M
replications/sec on 66 Gbps of aggregate network bandwith, comparable to the
same mark obtained by Cyclone on an aggregate 40 Gbps of network bandwidth when
not using the log on the SSD. We believe the general applicability afforded by
our software only solution together with the fact that we provide persistence by
design makes it a compelling alternative to an FPGA based solution even given
the larger latency.

@@

Using non-volatile memory to cache writes to flash is also employed in the
internal architecture of flash SSDs - including the SSD we use in the evaluation
of the paper. This ``write cache'' however does not help with small appends to
the write ahead log, as experiments with Rocksdb's own write-ahead log has shown
in this paper.

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request is currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores. This is along the lines of similar work~\cite{raft_lin}, using a fixed
client space and maintaining information about last seen sequence numbers from
clients in NVM. Distributed transactions on Cyclone shards however are outside
the scope of this paper.
